### 2017-02-03 paper/paper.tex
paper updates on friday

\documentclass[12pt,titlepage]{article}

% \usepackages21{fancyhdr}

\usepackage[printwatermark]{xwatermark}

\usepackage{grffile}
\usepackage{xcolor}
\usepackage{lipsum}
\usepackage{times}
\usepackage{soul}
\usepackage{epsfig}
\usepackage{rotating}
\usepackage{url}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{longtable}
\usepackage{keyval}
\usepackage{caption,subcaption}
\usepackage{arydshln}
%%\usepackage[hyphenbreaks]{breakurl}

%% allow urls to get broken on hyphens
\usepackage{hyperref}
\def\UrlBreaks{\do\/\do-}

%% APSR submission: no commas in citations between name and year
%% See http://merkel.zoneo.net/Latex/natbib.php
\bibpunct{(}{)}{;}{author-year}{}{;}

% the opening bracket symbol, default = (
% the closing bracket symbol, default = )
% the punctuation between multiple citations, default = ;
% the letter `n' for numerical style, or `s' for numerical superscript style, any other letter for author-year, default = author-year;
% the punctuation that comes between the author names and the year
% the punctuation that comes between years or numbers when common author lists are suppressed (default = ,);

\usepackage{footmisc}
\renewcommand{\footnotelayout}{\doublespacing} % set spacing in footnotes
\newlength{\myfootnotesep}
\setlength{\myfootnotesep}{\baselineskip}
\addtolength{\myfootnotesep}{-\footnotesep}
\setlength{\footnotesep}{\myfootnotesep} % set spacing between footnotes

% make footnote font size same as regular font size in text
\renewcommand{\footnotesize}{\normalsize}

%% Use this for a "DRAFT" watermark
%% \newwatermark[allpages,color=pink!50,angle=45,scale=5,xpos=-25,ypos=40]{DRAFT}

%% List all locations for graphics here
\graphicspath{ {../plots/} }

\begin{document}
\sloppy
\thispagestyle{empty}

%% APSR submission requires double-spaced footnotes
%%\newcommand{\footnote}[1]{\footnote{\doublespacing #1}} %% <-- note \doublespacing here.

\renewcommand{\topfraction}{.85}
\renewcommand{\bottomfraction}{.7}
\renewcommand{\textfraction}{.15}
\renewcommand{\floatpagefraction}{.66}
\renewcommand{\dbltopfraction}{.66}
\renewcommand{\dblfloatpagefraction}{.66}

\newcommand{\Yi}{\ensuremath{Y_i}}

% \urldef\myurlncsl1\url{foo%.com}
% \begin{document}
% text\footnote{WWW: \myurl}

\title{\Large{All in the family:\\Anna and Lisa Hahner's finishing
    times in\\the 2016 women's Olympic
  marathon}}\author{David Cottrell\thanks{Postdoctoral Research
  Fellow, Program in Quantitative Social Science, Dartmouth College,
    6108 Silsby Hall, Hanover, NH
    03755 (\texttt{david.cottrell@dartmouth.edu}).} \and Michael C.\
  Herron\thanks{Visiting Scholar, Hertie School of Governance, Berlin,
    Germany, and Professor of Government, Dartmouth College, 6108
    Silsby Hall, Hanover, NH 03755
    (\texttt{michael.c.herron@dartmouth.edu}).}}

\maketitle \doublespacing

%\begin{abstract} 
% \noindent
%The abstract
%\end{abstract}

%\newpage

\begin{quote}
  \emph{``I invested all I had and 300 meters before the finish line,
    I was next to Lisa. It was a magical moment that we could finish
    this marathon together. We did not think about what we were
    doing.'' -- Anna Hahner}
\end{quote}

\section*{Introduction}

At 9:30am on August 14, 2016, the Women's Olympic marathon kicked off
in Rio de Janeiro, Brazil, when 156 runners from 80 countries across
the world left the starting line en route to their destination 42.195
kilometers away. Two hours, twenty-four minutes, and four seconds
later, Jemima Sumgong of Kenya would be the first to cross the
finishline and take home gold; Sumgong was just three and one-half
minutes slower than her prior personal best time in the
marathon. Approximately 21 minutes later, twin marathoners from
Germany, Anna and Lisa Hahner, would cross the finishline together,
holding hands and celebrating a personal victory. Although the Hahners
would finish 81st and 82nd, respectively, well behind the winners of
the marathon, Anna Hahner would describe their joint finish as a
``magical moment.''

%% Honigkuchenpferde reference
%%
%% https://www.welt.de/sport/olympia/article157669264/Das-falsche-Laecheln-der-deutschen-Lauf-Zwillinge.html

The media quickly picked up on the Hanher story as an image of the
beaming twins finishing hand-in-hand captured a public audience. While
many believed the moment was a reflection of the Olympic spirit, not
everyone agreed with this rosy interpretation. The twins' happy facial
expressions at the finish were portrayed as a bit contrived---smiling
like ``Honigkuchenpferde,'' cookies in the shape of a horse---and the
sports director of the German Athletics Federation, Thomas
Kurschilgen, stirred up controversy when he suggested that the
Hahners' photo-finish was no coincidence.  Kurschilgen averred that
the twins slowed down so as to finish simultaneously and create a
spectacle which would ``generate media attention.'' Kurschilgen
justified his charge with the fact that the twins ran in the Rio
marathon at least 18 minutes slower their personal best times prior to
the
Olympics.\footnote{\url{https://www.nytimes.com/2016/08/17/sports/olympics/twins-finish-marathon-hand-in-hand-but-their-country-says-they-crossed-a-line.html}}
Not surprisingly, Kurschilgen's accusations were denied by the Hahner
twins, who claimed that their simultaneous finish was simply an
unintended coincidence.

%% Honigkuchenpferd quote from here:
%%
%% https://www.welt.de/sport/olympia/article157669264/Das-falsche-Laecheln-der-deutschen-Lauf-Zwillinge.html

What happened in the women's Olympic marathon, and how might we
develop a statistical approach that assesses whether the Hahner twin's
finish was coincidental or intentional?  These two interpretations are
clearly at odds. If the former, then the Hahners are to be celebrated
and their finish treated as an expression of the spirit behind the
Olympic games. If the latter, though, then the twins may have violated
this spirt by not trying hard enough. It is perhaps too easy for us to
write such a glib sentence---neither of us can fathom being able to
complete a marathon anywhere in the vicinity of two and a half
hours---but we nonetheless want to know what the data from the Olympic
marathon tell us.  Was the Hahner finish in the 2016 women's
Olympic marathon a lovely coincidence or something else?

Among female Olympic marathoners, the Hahner twins were not alone in
their familial ties.  The marathon also featured twins from North
Korea, Kim Hye-song and Kim Hye-gyong, who posted identical times and
finished 10th and 11th in the race, respectively. The Kim finish,
unlike the Hahner finish, appears devoid of post-race
controversy. Moreover, three triplets from Estonia competed in the Rio
marathon, although only two, Lily Luik and Leila Luik, finished it, in
97th and 114th place, respectively. The third Estonia triplet, Liina
Luik, recorded what is known as a DNF---an abbreviation that means did
not finish.  Although our focus here is the Hahner twins, we touch on
the Kim twins and Luik triplets in the course of what follows.

\section*{Marathon data and our research design}

For each participant who started the women's Olympic marathon, we know
several things: personal best marathon time prior to the 2016 Olympic
games; split times from the Rio marathon course at 5 kilometers, 10
kilometers, and so forth; and, finishing time. We cannot directly
observe the effort that an individual put into the race, and we do not
know why some runners have DNF results.  Some runners may have injured
themselves on the course and accordingly dropped out, and others may
have dropped out, uninjured, in anticipation of an unsatisfactory
result. Of the 156 marathon starters, 133 completed the race and 23
DNFed at various locations throughout the course. The overall DNF rate
was thus $\frac{23}{156} \approx 0.15$, and the relatively small
sample size at our disposal means that a 95\% confidence for this rate
is fairly wide, namely, $\left(0.098, 0.22\right)$.

The Kurschilgen accusation against the Hahner twins has two
components, that these two women ran slowly \emph{and} that they
finished simultaneously.  We suspect that Kurschilgen would not have
expressed ire at the Hahners had they finished in 1st and 2nd place in
Rio, hand-in-hand with wide grins, but of course we do not know this.
As such, our investigation of the charges that Kurschilgen offered
will distinguish between the idea of slow finish versus a simultaneous
finish.

Our research design is twofold.  First, we present visualizations that
describe various features of the Hahner twin's results, and an
important element in our visualizations is the difference between a
runner's Rio time and her prior personal best time in the marathon.
Overall, our visualizations suggest that the Hahner twins' pace in the
marathon was slow, but not excessively so, but that their simultaneous
finish was quite unusual given the twins' differences in abilities
(and similarly for the Kim twins).  We then turn to a regression-based
simulation of the marathon, which we develop based on relationships
(excluding the Hahner and Kim twins and Luik triplets) between known
personal best times and observed Rio finishing times.  Our simulations
show that the Hahner twins finished suspiciously close to each other
given the disparty between their personal best times.  Finally, we
return to Kurschilgen claims about the Hahners and offer our thoughts
about their validity.

\section*{Visualizing the Olympic marathon}

Figure \ref{fig:scatter} contains two plots, both of which describe
how finishing times from the Rio Olympic women's marathon varied as a
function of athletes' personal best times and their Rio half marathon
split times.  The points in the plots are colored by twin/triplet
status, and both plots contain second-order polynomial regression
lines.  We treat an athlete's prior personal best marathon time as a
measure of the athlete's underlying marathon talent, and we interpret
an athlete's half marathon split as a measure of the athlete's skill
at marathon running on August 14, 2016, the date of the women's
marathon in Rio.

\begin{figure}[!ht]
  \caption{Olympic results and measures of marathon ability}
  \label{fig:scatter}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot.pdf}
    \caption{Finishing and personal best times}
    \label{fig:45degreeplot}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot_half.pdf}
    \caption{Finishing and half marathon split times}
    \label{fig:45degreeplot_half}
  \end{subfigure}
\end{figure}

Figure \ref{fig:45degreeplot_half} includes a solid 45-degree line as
the axes in the figure correspond to full marathons; in addition, the
figure's rug marks denote personal best times of runners who DNFed.
Given the paucity of points (five of them) below the pictured
45-degree line, the vast majority of Olympic marathoners ran slower in
Rio compared to their personal bests.  Relative to personal best
times, the Hahner twins were definitely on the slow side, but a number
of runners had greater differences between their Olympic times and
their personal bests than the Hahners.  These runners are denoted with
squares in Figure \ref{fig:45degreeplot}, and there are 13 such
symbols, highlighting approximately 10\% of the racers who completed
the marathon.  The figure shows moreover that there were two women who
had personal best times slightly faster than the Hahner's and yet
finished after the German twins.  Although Figure
\ref{fig:45degreeplot} suggests that the Hahner twins were slower than
one would have expected given their previous best marathon times, it
is not consistent with the accusation that they dramatically slowed
down in the Rio marathon.

% > df <- read.csv("data/times.csv", stringsAsFactors = FALSE) %>% tbl_df()
% > df <- df[df$BIB != 1172,]
% > sort(dta$FINAL - dta$PB) / 60
%   [1] -2.3333333 -2.2166667 -2.0666667 -1.1166667 -0.4000000  0.2166667
%   [7]  0.6333333  0.9000000  1.0166667  1.3666667  1.5166667  1.7166667
%  [13]  1.7666667  1.8500000  2.1166667  2.4666667  2.5333333  3.0166667
%  [19]  3.0500000  3.1666667  3.2666667  3.2833333  3.3666667  3.4500000
%  [25]  3.6833333  3.9000000  3.9833333  4.2000000  4.2666667  4.3833333
%  [31]  4.4666667  4.5833333  4.6166667  4.6333333  4.6833333  4.7333333
%  [37]  4.7333333  4.8333333  5.0666667  5.0833333  5.0833333  5.1000000
%  [43]  5.1333333  5.3500000  5.3500000  5.4000000  5.7166667  5.7333333
%  [49]  5.9333333  6.0000000  6.0333333  6.1166667  6.3333333  6.3500000
%  [55]  6.5000000  6.5333333  6.6000000  6.6666667  6.7166667  6.7833333
%  [61]  6.8166667  6.9500000  7.1000000  7.1500000  7.2000000  7.3500000
%  [67]  7.4500000  7.5666667  7.6000000  7.6000000  7.6333333  7.7500000
%  [73]  7.8833333  7.9666667  7.9833333  8.3000000  8.6000000  9.3000000
%  [79]  9.5500000  9.5833333  9.6000000  9.7166667  9.9500000  9.9833333
%  [85] 10.3166667 10.5333333 10.6666667 10.7000000 10.8333333 10.9666667
%  [91] 11.1500000 11.1833333 11.4666667 11.7166667 12.1666667 12.6000000
%  [97] 12.9166667 13.2166667 13.2833333 13.3833333 13.7666667 13.8666667
% [103] 13.9166667 14.0500000 14.0833333 14.5666667 14.6833333 14.7500000
% [109] 15.0000000 15.0500000 15.5000000 15.9000000 16.2000000 16.9000000
% [115] 17.4500000 18.1000000 18.6666667 18.7166667 18.8000000 18.9333333
% [121] 20.1500000 20.6333333 20.7000000 20.8500000 21.3500000 23.4166667
% [127] 24.9333333 25.6500000 26.5666667 26.7666667 29.2166667 30.7000000
% [133] 30.9666667
% >

One limitation of using personal best times as indicators of
underlying athletic talent is that these times are potentially
confounded by the marathon courses at which they were set (some
courses, like the Berlin marathon, are known for fast times) and race
conditions like weather.  In addition, personal best times may not
capture raceday idiosyncrasies that might affect individual runners.
With that in mind, Figure \ref{fig:45degreeplot_half} plots marathon
finishing times against half marathon splits.  This figure has a
regression line as before but no 45-degree line.

Figure \ref{fig:45degreeplot_half} shows that there was nothing
abnormal about the Hahner twins' overall finishing times, conditional
on their half marathon split times.  This is consistent with the
previous Figure \ref{fig:45degreeplot} and inconsistent with
accusations made against the Hahner twins, at least the part of the
accusation that focused on their  overall pace.

If the Hahner twins did not slow down excessively, might they have
run somewhat strategically at the end of the race in order to generate
a simultaneous finish?  We now offer a visualization that speaks to
this question.

The personal best times of the Hahner twins were 115 seconds apart and
their official finishing times were separated by one second.  Is such
a 115 to one compression typical among pairs of runners or, in
contrast, was it unusual?  Are there other pairs of marathoners who
had a difference between personal best times of 115 seconds apart and,
if so, how close where their finishing times?

Of the 133 marathon finishers, there are $\binom{133}{2} = 8,778$
pairs of runners.  Of these and ignoring the Hahner twins, ten had
exactly a 115 second gap in personal best times.  Of these ten pairs,
differences in finishing times, in seconds, are as follows: 36, 93, 172,
319, 379, 459, 552, 671, 675, and 739.  In other words, of all pairs
of runners in the Rio marathon who had a personal best difference that
was equivalent to the Hahner twins' difference, the twins had the
greatest compression based on finishing time.

% df2b[abs(df2b$pb_diff)==115,]
% # A tibble: 11 × 8
%                       name_i              name_j pb_diff result_diff
%                        <chr>               <chr>   <int>       <int>
% 1                Anna Hahner         Lisa Hahner    -115          -1
% 2                  Lily Luik Mariya Korobitskaya     115          36
% 3        Lisa Jane Weightman      Lanni Marchant    -115          93
% 4           Irina Smolnikova      Viviana Chávez     115        -172
% 5              Sonia Samuels    Helalia Johannes     115        -319
% 6  Jessica Draskau-Petersson       Gladys Tejeda     115         379
% 7        Maryna Damantsevich       Gladys Tejeda     115         459
% 8          Panayióta Vlaháki     Matea Matosevic    -115         552
% 9         Nataliya Lehonkova       Olha Kotovska     115         671
% 10               Anja Scherl      Desireé Linden     115         675
% 11              Rose Chelimo    Helalia Johannes    -115        -739
% # ... with 4 more variables: diff_in_diff <int>, consecutive <lgl>,
% #   twins <chr>, PERCENTILE <dbl>
% >

\begin{figure}[!ht]
 \caption{Differences in personal best times and finishing times}
 \label{fig:diffdiffscatter}
 \centering
 \begin{subfigure}{.5\textwidth}
 \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_scatter_plot.pdf}
 \end{subfigure}
\end{figure}

We can generalize this result by looking at all pairs of runners in
the marathon.  For all 8,778 pairs of 133 finishers, Figure
\ref{fig:diffdiffscatter} plots differences in finishing times against
differences in personal best times, and pairs of twins/triplets are
identified by the same color scheme we used earlier.

Consider first the Hahner twins.  They are effectively located on the
horizontal axis because their difference in finishing times is one
second.  However, there are many points about the Hahner's orange dot,
and this shows that, conditional on an approximate 115 second
difference in personal best times, most marathoners did not have close
finishing times like the Hahnes.  Some pairs of runners with around
115 second personal best differences had finishing time differences of
1000 seconds, i.e., in excess of 15 minutes.  Of course the points in
Figure \ref{fig:diffdiffscatter} are not independent, but they provide
a sense of the dispersion in finishing time differences that one can
expect conditional on differences in personal best times.

Thinking about the accusations leveled against the Hahner twins,
Figure \ref{fig:diffdiffscatter} suggests that Anna and Lisa Hahner
did indeed run with an eye on each other.  In fact, the same can be
said of the Kim twins, who ran seemingly in lockstep throughout the
entire Rio marathon.  The North Korean twins had a personal best
difference of 53 seconds and a finishing time difference of literally
zero seconds.  Beyond these twins, there were eight pairs of Rio
runners with a 53 second personal best difference, and resulting
finishing time differences are as follows: 9, 51, 228, 340, 352, 571,
662, and 751.  As in the Hahner case, the Kim twins compressed their
finishing times more than any other pair of runners with a similar
personal best difference.

To get a sense of the lockstep nature of the Kim twins race, Figure
\ref{fig:secondsbehind} describes each runner's status at the various
split times on the marathon course.  Each dot in the figure depicts a
recorded split and the number of seconds each runner was behind the
race leader at the time.  There are more dots at earlier splits due to
the accumulation of DNFs.  The usual color scheme applies here, and
the Estonian DNF prior to the half marathon split is evident.  The
orange dots in the figure represent the Hahner twins, and as noted in
the figure Lisa Hahner was ahead of her sister through 15 kilometers
at which point Anna surged.  Figure \ref{fig:secondsbehind} contains
two red dots representing the North Korean Kim twins, but this is not
visually apparent because the Kim twins had identical split times
during the entire race marathon.

\begin{figure}[!ht]
  \centering
  \caption{Runner status by split}
  \label{fig:secondsbehind}
  \includegraphics[scale = 1]{seconds-behind.pdf}
  \begin{flushleft}
    \emph{Note: each dot represents one runner at a split.  DNFs are
      not pictured, and splits are not to scale.}
  \end{flushleft}
\end{figure}

\section*{Simulating the Marathon}

% The Hahner sisters claimed that they ran the marathon independent of
% each other.  Their simultaneous finish was no more than a product of
% ability and luck.  One might expect such a finish given that their
% personal best marathons are within 114 seconds of each other.  Add a
% little bit of luck to the occasion and, as the twins have claimed all
% along, one might easily argue that the simultaneous finish is simply a
% joyous coincidence rather than a wholly implausible event.

We now consider the following: if we take into consideration the
Hahner twin's similarities as runners as well as some natural
variation in finishing a marathon, what is the probability that the
twins finished at roughly the same time and/or sequentially?  To
answer this question we need to know the counterfactual distribution
of potential marathon finishing times that would have occurred had the
Hahner twins independently and repeatedly run the Rio marathon over
again.  Holding the twin's abilities constant, this would establish
the set of potential outcomes that would have occurred as result of
natural variation in the finishing the marathon.  We could use this
distribution of outcomes to determine the likelihood that a
simultaneous finish by the Hanhers, or at least a near-simultaneous
finish, occurred by chance alone.  If the twins rarely finish the
marathon together in such a counterfactual world, then one might be
skeptical that their observed finish occurred without
coordination---as they in fact claimed.

Unfortunately for us but nonetheless fortunately for the race's
participants, it is not possible to rerun the women's Olympic marathon
in Rio to establish a counterfactual distribution of potential race
outcomes for the Hahner twins.  However, we can attempt to simulate
this distribution by estimating the distribution of every other
runner's finishing time, conditioned on prior abilities, and drawing
from that distribution in order to observe the likelihood that Lisa
and Anna Hahner finish simultaneously.

To estimate the conditional distribution of Rio runner's result, we
assume that each runner's marathon time \Yi\ is distributed normally
with a mean that is a quadratic function of her personal best time
$X_i$:

$$\Yi \mid X_i \sim N\left(\beta_0 + \beta_1\,X_{i} + \beta_2\,X^2_{i}  + e_i, \sigma\right).$$
%%
where $N\left(\cdot,\cdot\right)$ denotes a normal distribution. We
estimate $\beta_1$ , $\beta_2$, and $\sigma$ using ordinary least
squares and our set of 133 finishing runners.  For a simulated
marathon, we then draw a runner's time from an estimated distribution,
conditioned on the runner's personal best outcome.  Once a race is
simulated, we locate Anna and Lisa Hahner and record both the time
between their finishes and the difference in their ranks.  We then
simulate a new race---drawing a new set of finishing times---and
record the same quantities.  In particular, the steps of the
simulation are as follows.

\begin{enumerate}
\item Ignoring twins and triplets, estimate with least squares a
  linear model that predicts a runner's finishing time ($Y_i$) based
  on her pre-Olympic personal best time ($X_i$) and her best time
  squared.
\item Extract coefficient vector, estimated covariance matrix, and
  estimated regression variance from this model.
\item For each simulated rate, draw intercept and slope estimates
  $\tilde{\beta}_0$, $\tilde{\beta}_1$, and $\tilde{\beta}_2$,
  respectively, from a multivariate normal distribution with mean
  equal to the estimated coefficient vector and covariance equal to
  the estimated covariance matrix.
\item For each runner, draw an error $\tilde{e}$ from a normal
  distribution with mean zero and a standard deviation equal to the
  standard deviation of the original regression model's residuals.
\item For each runner, predict the final result by combining the randomly
  generated beta coefficients and error terms, $\tilde{y} = \tilde{\beta}_0 +
  \tilde{\beta}_1\,X_i + \tilde{\beta}_2\,X_i^2 + \tilde{e}$.
\item Eliminate each runner from the race with a probability equal to
  the observed fraction of marathoners who did not finish the
  marathon.
\item Calculate the difference in time and difference in ranking
  between Anna Hahner and Lisa Hahner, assuming both finished the
  race.
\item Repeat above steps 10,000 times.
\end{enumerate}

Before considering results, DNF biz...

Figure \ref{fig:simdiff} contains two histograms of differences
between Anna and Lisa's finishes.  Figure
\ref{fig:simulatedfinishtimes} shows the distribution of absolute
differences in time where differences are grouped in 30 second bins
along the horizontal axis of the histogram.  The count for each bin is
denoted by the vertical length of the bars.  Figure
\ref{fig:simulatedranks} shows the distribution of the absolute
differences in Hahner twin rankings.  Differences are grouped as
single units ranging from no runner between the twins to nearly 120
runners between them.

\begin{figure}[!ht]
  \centering
  \caption{Distribution of simulated results for the Hahner twins}
  \label{fig:simdiff}
  \begin{subfigure}{.45\textwidth}
    \includegraphics[width=\textwidth,
    keepaspectratio]{simulated_time.pdf}
    \caption{Simulated differences in finishing times}
    \label{fig:simulatedfinishtimes}
  \end{subfigure}
  \begin{subfigure}{.45\textwidth}
    \includegraphics[width=\textwidth, keepaspectratio]{simulated_rank.pdf}
    \caption{Simulated differences in places}
    \label{fig:simulatedranks}
  \end{subfigure}
\end{figure}

The results plotted in the histograms question the credibility of Anna
and Lisa's story.  The results suggest that a simultaneous finish
would be very rare if Anna and Lisa had run independently of each
other.  For example, in less than 300 of the 10,000 simulated races
did Anna and Lisa finished within 30 seconds of each other .  And they
finished in consecutive rank in less than 175 of the 10,000 races.
Therefore, the close finish that we observed, where Anna and Lisa
crossed the finish line one-after-the-other, would be very unlikely if
we were to believe the twins raced independently of each other.

The results holds even if you replace their personal best time with
their time halfway through the race.  While this reduces the variation
in the predicted outcome of each runner and, therefore, reduces the
expected distance between Anna and Lisa, it is still quite rare for
Anna and Lisa to finish simultaneously or consecutively.

\begin{figure}[!ht]
  \centering
  \caption{Distribution of simulated results for the Hahner twins (using halfway split as predictor)}
  \label{fig:simdiffhalf} 
  \begin{subfigure}{.45\textwidth}
    \includegraphics[width=\textwidth,
    keepaspectratio]{simulated_time_half.pdf}
    \caption{Simulated differences in finishing times}
    \label{fig:simulatedfinishtimes}
  \end{subfigure}
  \begin{subfigure}{.45\textwidth}
    \includegraphics[width=\textwidth, keepaspectratio]{simulated_rank_half.pdf}
    \caption{Simulated differences in places}
    \label{fig:simulatedranks}
  \end{subfigure}
\end{figure}

% \section*{The probability of an unintentional simultaneous finish}

% Thomas Kurschilgen's accusations of an intentional finish stems from
% two observations. First, the Hahner twins finished slower than he
% expected given respective personal bests. Second, the twins finished
% together at the exactly same time. Kurschilgen clearly believed that
% neither of these events would have occurred had both Hahner run the
% marathon independently, absent coordination. By his logic, the German
% twins should have run faster and not have finished simultaneously.

% On the other hand, only a handful of runners completed the Rio
% marathon with times that were faster than their recorded best times.
% Thus, 18 minutes behind a personal record may not be the outlier that
% Kurschilgen claimed it to be. Moreover, there is good reason to
% suspect that an unintended simultaneous finish was more likely for the
% Hahner twin than it would be for any other set of runners. After all,
% these two women are twins with presumably similar abilities. Not only
% do they train together, but their pre-Rio best times are less than two
% minutes apart. While Anna may be slightly faster than Lisa measured by
% personal bests, we might expect the difference in their Rio result to
% be just as close as the differences in their recorded bests. And given
% random variation in finishing times, a simultaneous finish might not
% be out of the ordinary.

% Even though they seem similar, finishing at very similar times and
% finishing together are different phenomena. If, for example, both
% Hahners were of similar ability to each other and also to many other
% runners, then we might expect similar finishing times yet not
% necessarily similar placements. The latter will be a function of the
% extent to which all runners on the Rio course have similar talent
% levels.  This point is an important one and will be evident in the
% results that follow.

% To test the assertion that the Hahner twins paced themselves to finish
% at the same time and with back-to-back placements, we need to compute
% the probability that such a result would have occurred
% unintentionally. Hence we need to know the probability distribution of
% Anna and Lisa's finishing times if their runs had been independent of
% each other. The challenge, of course, is that this distribution is
% unknown.

% \section*{Modeling}

% The Hahner twins either acted independently $I$ or coordinated their
% times $C$; these are the two possible states of the race
% $\psi = \{I, C\}$. Given that we observed Anna and Lisa's finishing
% times, $Y_A = y_A$ and $Y_L = y_L$, we want to know the probability
% that they ran independently, as they say they did. Hence, we are
% looking to determine,

% $$P(\psi = I \mid y_A  \cap y_L )$$

% However, to determine this, we need to have some understanding of a
% likelihood function that specifies Rio finishing times. We want to
% know the likelihood of Anna's and Lisa's final times given
% independence $P(y_A \cap y_L \mid \psi = I )$. We can estimate this
% function with a few assumptions. First, we assume that under
% independence, any given runner's final time $Y_i$ is conditional on
% his/her running ability plus noise. Specifically, we assume that the
% $Y_i$ is a linear function of the runner's ability $X_i$ plus a
% normally distributed error term $e_i \sim N(0, \sigma)$. Second, we
% assume that every runner shares the same linear relationship - meaning
% the slope and intercept remain constant across runners. Third, we
% assume that the error term is drawn from a common distribution across
% runners. Hence, luck and misfortune are drawn from the same
% distribution. Therefore,

% $$Y_i \sim N(\beta_0 + X_{i}\beta_1 + e_i, \sigma)$$

% We also assume that a runner's ability $X_i$ can be measured precisely
% by her best marathon performance leading up to the Olympics.

% %% Any measurement error must therefore be negligible.

% If Anna and Lisa intentionally slowed down as a result of coordination
% then we would likely observe $y_A > E(Y_A \mid \psi = I) $ and
% $y_L > E(Y_L\mid \psi = I )$. In other words, the final times that
% Anna and Lisa recorded in the race would be greater than we would
% expect if they had run independently.

% Moreover, if Anna and Lisa coordinated to finish simultaneously, then
% the difference between the two sisters' final times would be less that
% the expected difference had they run independently. Therefore, under a
% coordinated finish we would expect $\left|y_A - y_L\right| < E(\left|Y_A - Y_L\right| \mid \psi = I )$.

% \section*{Did Anna and Lisa intentionally slow down?}

% According to Kurschilgen, Anna and Lisa underperformed in the Rio marathon. He claimed that because their goal was to finish simultaneously rather than finish at their fastest pace, their times were slower than they otherwise would have been.  He claimed that the twins were simply trading speed for a photo-finish.

% If this were the case, the function generating Anna and Lisa's final times would deviate from the function that generated everyone else's final times.  Given everyone else would draw their times from the independent distribution  $Y_i \sim N(\beta_0 + X_{i}\beta_1 + e, \sigma)$,  Anna and Lisa would draw from a distribution of times that are slower in expectation. Hence, they would lie well-above the line that links a runner's performance in Rio to their previous best performance.

% \begin{figure}[!ht]
%   \caption{Relationship between Personal Best and Result  -- caption?}
%   \label{fig:scatter}
%   \begin{subfigure}{.5\textwidth}
%     \centering
%     \includegraphics[width=\textwidth, keepaspectratio]{scatter_plot.pdf}
%     \caption{Rio finishing times and personal best times}
%     \label{fig:45degreeplot}
%   \end{subfigure}
%   \begin{subfigure}{.5\textwidth}
%     \centering
%     \includegraphics[width=.975\textwidth, keepaspectratio]{studentized_residuals.pdf}
%     \caption{Studentized residuals}
%     \label{fig:studentizedresiduals}
%   \end{subfigure}
% \end{figure}

% Figure \ref{fig:45degreeplot} displays the relationship between a
% runner's personal best (horizontal axis) and her final result
% (vertical axis).  The 133 finishers are plotted as grey points, and
% the 23 DNFs are depicted below along the horizontal axis with
% tick-marks.  The dashed black line with grey confidence intervals
% represent ordinarily least squares estimates of the linear
% relationship between racer's finishing and personal best times.

% One can see in Figure \ref{fig:scatter} that there is a clear
% relationship between a runner's personal best time and her performance
% in the Olympics.  While there is a significant amount of noise in the
% relationship, the Hahner twins' dots are well above the regression's
% fitted line notwithstanding this point.  They finished the Rio
% marathon much slower than they should have, per their pre-Olympic
% personal best times.  In fact, if we sort studentized residual, we can
% see in Figure \ref{fig:studentizedresiduals} that the Hahner twins are
% in the tail-end of the residual distribution.  Anna Hahner's residual
% deviation is greater than 91.7\% of the runners who completed the
% marathon while Lisa Hahner's deviation is greater than 86.5\%.  Hence,
% the twins appear to have finished at a much slower pace than expected,
% which is what we might expect if they were coordinating their runs.
% The North Korean Kim twins are also in the tail of the residual
% distribution---but the left tail.  These twins finished much faster
% than expected.

% \section*{Did Anna and Lisa intentionally finish together?}

% Anna and Lisa finished the marathon at an unusually slow rate, which is what we might expect if the twins intentionally slowed their pace to finish simultaneously.  However, the observation of a slow finish does not necessarily imply that it was an intentional result .  It only suggests that Kurschilgen's claim regarding this issue holds up in the data.

% Perhaps it would be easier to refute the claim if we could show that the Twins's back-to-back finish was not such an unusual event.  Given that Anna and Lisa were so close in their personal best times, it may seem reasonable to expect that they would also be close in their finishing times.   Hence, one way to determine if the twins's finish was intentional is to quantify the likelihood that such a close finish would occur by chance given the differences in their ability.

% We can attempt to quantify this likelihood by looking at how differences in the personal best time of two runners translate into differences in Olympic times .  If Anna and Lisa's finish was unintentional we'd expect that such a close finish would be unlikely given their difference in personal best times.   Yet, how does the difference in personal best times relate to the difference in finishing times?

% One way we can estimate this relationship is to look at all of the differences that we observe in the marathon.  We can take the difference in personal best time for every combination of runners in the data and look at the difference in their final times.  We should expect that compared to the difference in their personal best times, the difference in Anna and Lisa's final times should be commensurate with the conditional difference of all other runners. %This needs a better explanation.

% Therefore, we find all combinations of runners in the marathon and calculate two quantities: 1) the difference in their personal best time and 2) the difference in their final time. We plot the relationship between the two quantities in Figure \ref{fig:diffdiffscatter}.   All 8,778 dyadic combinations of runners are displayed by grey points in the scatter plot.   The solid black line indicates a one-to-one relationship, where every point would fall if the difference in final times were systematically equivalent to the difference in personal bests.

% % \begin{figure}[!ht]
% %  \caption{Relationship between Difference in Personal Best and Difference in Result}
% %  \label{fig:diffdiffscatter}
% %  \centering
% %  \begin{subfigure}{.5\textwidth}
% %  \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_scatter_plot.pdf}
% %  \end{subfigure}
% % \end{figure}

% We can see from the yellow dot that the Hahner Twins are just below this one-to-one line.  Hence, the twins' final times were closer to each other than their personal best times; but not by much.  In fact, this might be an expected difference if we expected there to be a one-to-one relationship.     However, although there is a clear relationship between the difference in final time and the difference in personal best, it is not a strong one-to-one relationship.  We've added a dotted smoother to display a cubic splines estimate of the conditional averages.  While the Hahner twins may be close to the one-to-one line, they are much further away from the average set of runners.  Hence, we would expect for any pair of runners, the difference in their final results should be much greater than their difference in their personal best.  However, the Hahner twins do not fit this expectation.   Yet, how different from this expectation are they?

% One way we can quantify the degree to which the twins are outliers is by comparing their difference in difference to every other pair of runner's difference-in-differences.  We calculate the difference-in-differences for each pair of runners by simply subtracting their difference in result (y-axis) from their difference in personal best (x-axis).  Therefore positive difference-in-differences mean that the runners ended up further apart than their difference in personal bests.  Negative difference-in-differences mean that the runners end up closer together.  Hence, the Hahner twins have a difference-in-difference of -114 seconds.  They finished 114 seconds closer together than the distance between their personal bests.

% We arrange all 8,778 difference-in-differences from small to large in the first plot in Figure \ref{fig:diffdiff}. The Hahner twins are denoted by the yellow point.  Using this plot we can visualize where the twins fall within the distribution of difference-in-differences.  It is clear that they are closer to the lower tail of the distribution.  However, compared to all other difference-in-differences, their result is not that extreme.  Nearly 21\% of the difference-in-dfferences were less than the that of the twins'. Hence, we'd expect at least one in five pairs of runners to have a reduction in difference at least as great as the twins.

% However, if we limit the observations to only those runners who had personal best times that were at least as close to each other as the Hahner twins, we get a different perspective.  Among those runners,  the Hahner twins reduced the distance between them to the greatest extent.  There was no other pair of runners whose reduced the difference in the final time relative to their difference in personal bests more than them.

% \begin{figure}[!ht]
%  \caption{How Hahner Twins Rank in Difference-in-Differences}
%  \label{fig:diffdiff}
%  \begin{subfigure}{.5\textwidth}
%  \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_1.pdf}
%  \end{subfigure}
%  \begin{subfigure}{.5\textwidth}
%  \includegraphics[width=\textwidth, keepaspectratio]{diff_in_diff_2.pdf}
%  \end{subfigure}
% \end{figure}

\section*{Conclusion}

\newpage
\section*{Further reading}

\begin{itemize}

\item \url{http://www.telegraph.co.uk/olympics/2016/08/17/german-twins-criticised-for-finishing-olympic-marathon-fun-run-h}

\item
  \url{https://www.nytimes.com/2016/08/17/sports/olympics/twins-finish-marathon-hand-in-hand-but-their-country-says-they-crossed-a-line.html}

\item
  \url{https://www.welt.de/sport/olympia/article157669264/Das-falsche-Laecheln-der-deutschen-Lauf-Zwillinge.html}
\end{itemize}

\end{document}