### 2017-02-17 paper/chance_memo.tex
ref memo outline

\documentclass[12pt]{article}
\sloppy
\usepackage{times}
\usepackage{url}
\usepackage{arydshln}

%% uncomment to use a watermark
%\usepackage{draftwatermark}
%\SetWatermarkScale{1}

\begin{document}
\thispagestyle{empty}

\noindent {\Huge Memo}\\[0.5in]

\noindent
\begin{tabular}{ll}
  \textbf{To:} &  Scott Evans, editor, \emph{Chance}\\
  \textbf{From:} & David Cottrell and Michael C.\ Herron, Dartmouth College\\
  \textbf{Date:} & \today \\
  \textbf{Re:} &  Revisions to women's marathon submission
\end{tabular}

\vspace{.1in}
\hrule
\vspace{.2in}

\bigskip

This memorandum is a response to your February 16, 2017, email
regarding our manuscript, ``All in the family: German twin finishing
times in the 2016 women's Olympic marathon.'' Thank you very much for
inviting us to submit a revision of our manuscript, and this
memorandum explains how we have updated it in light of comments from
you and the Chance Editorial Board.

blah blah blah

 \end{document}